---
title: "Average data before/after COVID"
output: html_document
---

In this section, we will take a look at the average export/import values and volumes of the United States by product type before and after Covid-19 pandemic using bar graphs. In this analysis, we defined before Covid-19 as from 2018 to 2019 and after Covid-19 as from 2020 to 2021.

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(dplyr)
library(plotly)
library(ggpubr)

theme_set(theme_minimal())

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r load data, include=FALSE}
import_value_df = 
  read_csv("data/cleaned data/import_value_combined.csv") %>% 
  mutate(
    product_type = recode(product_type, "Vegetables & preparations" = "Vegetables and preparations","Fruits & preparations" = "Fruits and preparations", "Grains & feeds" = "Grains and feeds","Beef & veal" = "Beef and veal", "Coffee & products" = "Coffee and products", "Cotton & linters" = "Cotton and linters","Cut flowers & nursery stock" = "Cut flowers and nursery stock","Oilseeds & products" = "Oilseeds and products","Red meats & products" = "Red meats and products", "Sugar & related products" = "Sugar and related products","Tree nuts & preparations" = "Tree nuts and preparations","Cocoa & products" = "Cocoa and products","Cattle & calves" = "Cattle and calves")
  )

export_value_df = 
  read_csv("data/cleaned data/export_value_combined.csv") %>% 
  mutate(
    product_type = recode(product_type,"Cotton & linters" = "Cotton and linters","Red meats & products" = "Red meats and products","Tree nuts & preparations" = "Tree nuts and preparations","Grains & feeds" = "Grains and feeds", "Hides & skins" = "Hides and skins","Poultry meats & products" = "Poultry meats and products","Fruits & products" = "Fruits and products","Sugar & tropical products" = "Sugar and tropical products","Animal feeds & oil meal" = "Animal feeds and oil meal", "Vegetables & products" = "Vegetables and products")
  )

import_volume_df = 
  read_csv("data/cleaned data/import_volume_combined.csv") %>% 
  mutate(
    product_type = recode(product_type, "Vegetables & preparations" ="Vegetables and preparations","Fruits & preparations" = "Fruits and preparations", "Grains & feeds" = "Grains and feeds","Beef & veal" = "Beef and veal", "Coffee & products" = "Coffee and products", "Cotton & linters" =  "Cotton and linters","Cut flowers & nursery stock" = "Cut flowers and nursery stock","Oilseeds & products" = "Oilseeds and products","Red meats & products" = "Red meats and products", "Sugar & related products" = "Sugar and related products","Tree nuts & preparations" = "Tree nuts and preparations","Cocoa & products" = "Cocoa and products", "Cattle & calves" = "Cattle and calves")
  )

export_volume_df = 
  read_csv("data/cleaned data/export_volume_combined.csv") %>% 
  mutate(
    product_type = recode(product_type,"Cotton & linters" = "Cotton and linters","Red meats & products" = "Red meats and products","Tree nuts & preparations" = "Tree nuts and preparations","Grains & feeds" = "Grains and feeds", "Hides & skins" = "Hides and skins","Poultry meats & products" = "Poultry meats and products","Fruits & products" = "Fruits and products","Sugar & tropical products" = "Sugar and tropical products","Animal feeds & oil meal" = "Animal feeds and oil meal")
  )
```


```{r message=FALSE}
before_covid_import_value =
  import_value_df %>% 
  filter(year == "2018" | year == "2019") %>% 
  add_column("Covid_19" = "before") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(import_value)) 


after_covid_import_value =
  import_value_df %>% 
  filter(year == "2020" | year == "2021") %>% 
  add_column("Covid_19" = "after") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(import_value)) 

covid_import_value_df = 
  rbind(before_covid_import_value,after_covid_import_value)

value_pct = 
  covid_import_value_df %>%
  group_by(product_type) %>% 
  mutate(after = lead(average)) %>%
  mutate(pct.change = (after - average) / average) %>% 
  arrange(desc(pct.change))

before_covid_export_value =
  export_value_df %>% 
  filter(year == "2018" | year == "2019") %>% 
  add_column("Covid_19" = "before") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(export_value))

after_covid_export_value =
  export_value_df %>% 
  filter(year == "2020" | year == "2021") %>% 
  add_column("Covid_19" = "after") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(export_value)) 

covid_export_value_df = 
  rbind(before_covid_export_value,after_covid_export_value)

```


### Average Import Value Before/After Covid From 2018 To 2021

The plot below shows the United States’ average import value by product type before and after Covid-19. In general, the average import values increased by approximately 16.53% after Covid-19 pandemic. Some of the products’ import value, such as fruits and preparations, Cotton and linters, beef and veal, coca and products, vegetable oils, oilseeds and products and red meats and products, have all increased by more than 20%. In addition, it is interesting to note that both distilled spirits and manufactured tobacco are only imported after the pandemic.

```{r}
covid_import_value_plot =  
  covid_import_value_df %>%
  ggplot(aes(x = product_type, y = average, fill = Covid_19)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Product Type",
    y = "Average Import Values"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)
    ) 

```

<center>
```{r}
ggplotly(covid_import_value_plot)
```
</center>


### Average Export Value Before/After Covid From 2018 To 2021

This plot shows the United States’ average export values by product type before and after Covid-19. In general, the average export values increased by approximately 18.44% after Covid-19 pandemic. Some of the products’ export value, such as corn, soybeans, dairy products, other oilseeds, and grains and feeds, have all increased by more than 20%. It is interesting to notice that ethanol, which is a renewable fuel made from corn and other plants, tobacco, distilled spirits, are only exported to other countries after the pandemic.

```{r}
covid_export_value_plot =  
  covid_export_value_df %>%
  ggplot(aes(x = product_type, y = average, fill = Covid_19)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Product Type",
    y = "Average Export Values"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)
    )


```

<center>
```{r}
ggplotly(covid_export_value_plot)
```
</center>


```{r message=FALSE}
before_covid_import_volume =
  import_volume_df %>%
  filter(year == "2018" | year == "2019") %>% 
  add_column("Covid_19" = "before") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(import_volume))

after_covid_import_volume =
  import_volume_df %>% 
  filter(year == "2020" | year == "2021") %>% 
  add_column("Covid_19" = "after") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(import_volume))

covid_import_volume_df = 
  rbind(before_covid_import_volume,after_covid_import_volume)

before_covid_export_volume =
  export_volume_df %>%
  filter(year == "2018" | year == "2019") %>% 
  add_column("Covid_19" = "before") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(export_volume)) 

after_covid_export_volume =
  export_volume_df %>% 
  filter(year == "2020" | year == "2021") %>% 
  add_column("Covid_19" = "after") %>% 
  group_by(product_type, Covid_19) %>% 
  summarize(average = mean(export_volume)) 

covid_export_volume_df = 
  rbind(before_covid_export_volume,after_covid_export_volume)

```

## Average Import Volume Before/After Covid From 2018 To 2021

The plot below shows the United States’ average import volume by product type before and after Covid-19. In general, the average import volume increased by approximately 3.62% after Covid-19 pandemic. Some of the products’ export volume, such as wine, beef and veal, and coca and products, have increased by approximately 10%. 

```{r}
covid_import_volume_plot =  
  covid_import_volume_df %>%
  ggplot(aes(x = product_type, y = average, fill = Covid_19)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Product Type",
    y = "Average Import Volumes"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)
    ) 


```

<center>
```{r}
ggplotly(covid_import_volume_plot)
```
</center>

## Average Export Volume Before/After Covid From 2018 To 2021

This plot shows the United States’ average export volume by product type before and after Covid-19. In general, the average export volume increased by approximately 9.23% after Covid-19 pandemic. Some of the products’ export volume, such as sorghum, nonfat dry milk, red meats and products and coarse grains, have all increased by more than 10%. It is worth to point out that the export volume of sorghum have increased by 91.6%.

```{r}
covid_export_volume_plot =  
  covid_export_volume_df %>%
  ggplot(aes(x = product_type, y = average, fill = Covid_19)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Product Type",
    y = "Average Export Volumes"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6)
    )

```

<center>
```{r}
ggplotly(covid_export_volume_plot)
```
</center>
